SSA based global code motion of identical computations

### 2017-04-16 acmart-pacmpl-template.tex
ackn

\newpage

@@

\section{Profitability check (Cost models)}
\label{sec:cost-models}
After the legality checks have passed, we check if a code-motion is profitable.
That takes into account the impact code-motion would have on various parameters
that might affect runtime performance e.g., impact on live-range, gain in the
code size. Since this is mostly a code-size optimization pass, the goal is to
not regress in performance across popular benchmarks at the same time reduce
code size as much as possible. Following are the cost models which are
implemented:

\subsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill), but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after code-motion is less than before it will
be moved.

@@

In this case, since `b' and `c' are killed in `a0' and `a1', hoisting them will
reduce the register pressure in B3 and B4 because two operands will be freed.

@@

\subsection{In the presence of calls}
\label{cost:across-calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored. So before the
call, the register pressure is high because the number of available registers
are reduced by the number of caller saved registers. In that situation if a
computation is hoisted across the call, that would increase the total number of
registers required by one, thus contributing to the register pressure. However,
in the special case discussed in Section~\ref{hoist:reg-pressure}, it will be
okay to hoist because the register pressure would not decrease.

Hoisting loads/stores across calls also require precise analysis of all the
memory addresses accessed by the call. Our implementation being an
intraprocedural pass, the analysis is very conservative. In the presence of pure
calls, loads can be hoisted but stores can't. Also, if the call throws
exceptions, or if it may not return, memory references cannot be hoisted.

@@

\newpage

\newpage

@@

\section{Acknowledgments}
We would like to thank Daniel Berlin for his code reviews and for his feedback
on earlier versions of this paper and Brian Grayson for motivating examples that
started this work.

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

@@

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

>>>>

@@

\subsection{Profitability check (Cost models)}
\label{sec:cost-models}
After the legality checks have passed, we check if a code-motion is profitable.
That takes into account the impact code-motion would have on various parameters
that might affect runtime performance e.g., impact on live-range, gain in the
code size. Since this is mostly a code-size optimization pass, the goal is to
not regress in performance across popular benchmarks at the same time reduce
code size as much as possible. Following are the cost models which are
implemented:

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill), but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after code-motion is less than before it will
be moved.

@@

In this case, since `b' and `c' are killed in `a0' and `a1', hoisting them will
reduce the register pressure in B3 and B4 because two registers will be freed.

@@

\subsubsection{Hoisting expression across call}
\label{cost:across-calls}
Even hoisting scalars across calls is tricky because it can increase the number
of spills. During the frame lowering of calls, the argument registers, in
general, the caller saved registers are saved because they might be modified by
the callee and after the call they are restored. So before the call, the
register pressure is high because the number of available registers are reduced
by the number of caller saved registers. In that situation if a computation
that increases register pressure is not profitable to hoist.

Hoisting loads/stores across calls also require precise analysis of all the
memory addresses accessed by the call. The current implementation being an
intraprocedural pass, cannot hoist aggressively across calls. In the presence of
pure calls, loads can be hoisted but stores can't. Also, if the call throws
exceptions, or if it may not return, nothing can be hoisted.

@@

@@

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
We would like to thank Daniel Berlin for his code reviews and for his feedback
on earlier versions of this paper and Brian Grayson for motivating examples that
started this work.
\end{acks}

@@

%% Appendix
%\appendix
%\section{Appendix}
%Text of appendix \ldots
